Firstly, I would like to give my sincere gratitude to my advisor Dr. Sarah E. Rothenberg, for her valuable advice, patient guidance, and continuous support and encouragement. Her research intuition and experiences inspired my keen interest in the field of environmental health over the course of my Ph.D. studies. I could not have imagined finishing my dissertation in such a short time without her help.

Also, I would like to thank the rest of my committee. My sincere thanks goes to Dr. Jihong Liu, for her professional suggestions and inspiration on this dissertation, as well as her friendship and help in making me feel comfortable in Columbia. I am also deeply grateful to Dr. Anindya Chanda and Dr. David C. Volz for giving me so many valuable comments and advice on my dissertation research. 

I would also like to express my appreciation to all my friends who have helped me at USC. I want to thank Dr. Chandrani Mitra, one of my best friends in the Department of Environmental Health Sciences, for her support and useful advice during my most difficult times. My special thanks will also be given to my roommates, Guangfang Li and Dr. Shengnan Meng, with whom I've made the warmest memories in Columbia.

Last but not least, I would like to give my heartfelt thanks to my dear parents. They are the ones who have always supported me and loved me unconditionally. I'm so proud to have them as my parents and also my best friends in life. I also want to thank my boyfriend, Dr. Shu Yan. He is always there when I need him. 
